---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
###Retrieve Data

The data are available at https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

Download and place in your working directory

```{r echo=TRUE}
  data<-read.csv("activity.csv")
```

###Examine the data
```{r echo=TRUE}
  summary(data)
  str(data)
  head(data)
  class(data)          # "data.frame"
  sapply(data, class)  # show classes of all columns
  typeof(data)         # "list"
  names(data)          # show list components
  dim(data)            # dimensions of object, if any

```

```{r echo=FALSE, results='hide'}
  library(dplyr)
  library(ggplot2)
  library(stringr)
  library(scales)
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
  by_day<-group_by(data,date)
  sum_by_day<- summarise(by_day,total=sum(steps))
  hist(sum_by_day$total)  
```

### Find Mean
```{r echo=TRUE}
  avg_by_day<- summarise(by_day,average=mean(steps))
  qplot(date,total,data=sum_by_day)
```

### Find Median
```{r echo=TRUE}  
  middle_by_day<-summarise(by_day,median=median(steps,na.rm=TRUE))
  qplot(date,median,data=middle_by_day)
```


## What is the average daily activity pattern?
###convert the interval to a valid time
```{r echo=TRUE}
  data$Time<-strptime(str_pad(as.character(data[,3]),4,side="left",pad="0"),"%H%M" )
  
```

###Put the date and time together
```{r echo=TRUE}
  digits4<-str_pad(as.character(data[,3]),4,side="left",pad="0")
  data$interval6<-paste0(substr(digits4,1,2),":",substr(digits4,3,4),":00")
  data$datetimeString<-paste(as.character(data[,2]), data$interval6)
  

df<-data[,c(1,5)]
df_grp<-group_by(df,interval6)
df_sum<-summarize(df_grp,total=mean(steps,na.rm=TRUE))
ggplot(df_sum,aes(interval6,total))+geom_point()
```

## Imputing missing values
###how many NA's are there
```{r echo=TRUE}
  length(which(is.na(data[1])==T))
```

###Make a new dataset with a ZERO where a NA was in original
```{r echo=TRUE}
  data2<-data[,c(1,2,3)]
  data2[is.na(data2[1,])==T,1] <- 0
```

###Histogram

```{r echo=TRUE}
  by_day<-group_by(data2,date)
  sum_by_day<- summarise(by_day,total=sum(steps))
  hist(sum_by_day$total)  
```

### Find Mean
```{r echo=TRUE}
  avg_by_day<- summarise(by_day,average=mean(steps))
  qplot(date,total,data=sum_by_day)
```

### Find Median
```{r echo=TRUE}  
  middle_by_day<-summarise(by_day,median=median(steps,na.rm=TRUE))
  qplot(date,median,data=middle_by_day)
```

## Are there differences in activity patterns between weekdays and weekends?

